---
title: "Spatial and spatiotemporal correlations in size from WCBTS"
author: "Eric Ward"
date: "3/28/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(sdmTMB)
library(GGally)
```

```{r}
# load grid
grid = readRDS("survey_data/wc_grid.rds")
grid$lon = grid$X * 10
grid$lat = grid$Y * 10
grid$depth_scaled = as.numeric(scale(-grid$depth))
grid$depth_scaled2 = grid$depth_scaled^2
# load spp
spp = readRDS("spp.rds")

```

### Synchrony in year effects  

There's a lot of interesting correlations here -- e.g. positive correlations among lingcod and a number of sole spp. Positive also among a few rockfishes. Time series are short however (~ 2003 - 2020).

```{r}
# extract year effects
for(i in 1:30) {
  m = readRDS(paste0("bio_runs/",i,".rds"))
  tm <- tidy(m,effects=c("fixed"), conf.int=TRUE)[-c(1:2),]
  tm$species <- as.character(spp$scientific_name[i])
  if(i == 1) {
    df = tm
  } else {
    df = rbind(df, tm)
  }
}
spp = dplyr::rename(spp, species=scientific_name)
df = dplyr::left_join(df, spp)# join in common name

wide_mat <- dplyr::filter(df, !is.na(estimate), estimate>0) %>% 
  dplyr::select(term, estimate, common_name) %>%
            tidyr::pivot_wider(names_from = common_name,
                  values_from = estimate)
#corMat <- cor(wide_mat[,-1], use="pairwise.complete.obs")

ggcorr(wide_mat[,-1])
```

\break  

### Magnitude of spatial correlation
```{r}
# extract ran pars
for(i in 1:30) {
  m = readRDS(paste0("bio_runs/",i,".rds"))
  tm <- tidy(m,effects=c("ran_pars"), conf.int=TRUE)
  tm$species <- as.character(spp$species[i])
  if(i == 1) {
    df = tm
  } else {
    df = rbind(df, tm)
  }
}

df = dplyr::filter(df, conf.high < 1000)
```

Like previous models, we see the magnitude of sigma_E being generally greater than sigma_O -- though in some of the rockfishes they're pretty similar in magnitude.    

```{r}
dplyr::filter(df, term %in% c("sigma_E","sigma_O")) %>%
ggplot(aes(species, estimate, col=term)) +
  geom_linerange(aes(ymin=conf.low,ymax=conf.high)) +
  geom_point()+
  coord_flip() + ylab("") +
  xlab("Estimate") + theme_bw()
```

The majority of species don't really support temporal autocorrelation in spatiotemporal fields. Though several (lingcod, longnose skate, chilipepper rockfish) have stronger positive correlation (rho). Bocaccio seems to be a bit of an outlier with negative autocorrelation ~ -1.0.  

```{r}
dplyr::filter(df, term %in% c("rho")) %>%
ggplot(aes(species, estimate, col=term)) +
  geom_linerange(aes(ymin=conf.low,ymax=conf.high)) +
  geom_point()+
  facet_wrap(~term,scale="free") + 
  coord_flip() + ylab("") +
  xlab("Estimate") + theme_bw()
```

\break   

### Are estimated spatial fields correlated?

For this, I'm projecting to the survey grid, and then asking if the estimated random fields are correlated -- seems to be some decent correlation spatially for some groups.

```{r}
grid$year <- min(m$data$year)
grid_yr = grid
grid_yr$year <- min(m$data$year)
for(y in 2004:max(m$data$year)) {
  tmp <- grid_yr
  tmp$year <- y
  grid = rbind(grid, tmp)
}

for(i in 1:30) {
  m = readRDS(paste0("bio_runs/",i,".rds"))
  pred = predict(m, newdata = dplyr::filter(grid, year %in% unique(m$data$year)))
  pred$species <- spp$common_name[i]
  if(i==1) {
    pred_all <- pred
  } else {
    pred_all = rbind(pred_all, pred)
  }
}

```


```{r}
spat_mat = 
  dplyr::filter(pred_all, year == 2018, omega_s>0) %>% 
  dplyr::select(omega_s, species, year, lon, lat) %>%
            tidyr::pivot_wider(names_from = species,
                  values_from = omega_s)

ggcorr(spat_mat[,-c(1:3)])
```

```{r fig.height=3}
corMat <- cor(spat_mat, use="pairwise.complete.obs")
hist(corMat[upper.tri(corMat)],40,main="",xlab="Pairwise correlations (omega_s)")
```

\break   

### Are estimated spatiotemporal fields correlated?

Same thing as with spatial field above, but spatiotemporal version. I think what we see here is generally weaker correlations in the spatiotemporal vs spatial fields, and fewer cases of extremes ()

```{r}
spat_mat = 
  dplyr::filter(pred_all, epsilon_st>0) %>% 
  dplyr::select(epsilon_st, species, year, lon, lat) %>%
            tidyr::pivot_wider(names_from = species,
                  values_from = epsilon_st)

ggcorr(spat_mat[,-c(1:3)])
```


```{r fig.height=3}
corMat <- cor(spat_mat, use="pairwise.complete.obs")
hist(corMat[upper.tri(corMat)],40,main="",xlab="Pairwise correlations (epsilon_st)")
```

